---
title: "fusionData"
author: Kevin Ummel ([ummel@sas.upenn.edu](mailto:ummel@sas.upenn.edu))
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE, # Useful for testing/editing; will create /README_cache directory
  comment = NA,
  fig.path = "man/figures/README-"
)
```

```{r, echo = FALSE, include = FALSE}
library(fusionData)
```

## Overview

**fusionData** is used to create and manage the data inputs underpinning the larger fusionACS platform. It facilitates a number of steps in the overall fusionACS workflow:

1. **Ingest**: Process raw survey data using a standard approach and formatting.
1. **Document**: Document survey variables and compile a "universal" data dictionary.
1. **Harmonize**: Harmonize variables in "donor" surveys with those in the American Community Survey (ACS).
1. **Compile spatial data**: Compile data from multiple spatial datasets for merging with survey microdata.
1. **Prepare for fusion**: Prepare donor and ACS/recipient microdata inputs for the [fusionModel package](https://github.com/ummel/fusionModel).

## Setup and install

The fusionData *master* branch can be cloned to its own project directory on your local machine using RStudio and [these instructions](https://book.cds101.com/using-rstudio-server-to-clone-a-github-repo-as-a-new-project.html).

Use the following setup parameters:
```r
Repository URL: https://github.com/ummel/fusionData
Project directory name: fusionData
```

You will be prompted to enter your Github username and password. After the repository has been cloned to your local project directory, you can install and load the package itself. You may be prompted to install some dependencies.

```r
# Install the fusionData package locally
devtools::install()

# Load fusionData package
library(fusionData)
```

For full functionality, it is necessary to download the remotely-stored processed survey microdata and processed spatial data files. The following section ("Structure and usage") has more detail about this and the associated reasoning. See `?getSurveyProcessed` and `?getGeoProcessed` as well. To get up and running, you'll need to call the following commands.

```r
# Download all remote processed survey microdata files
getSurveyProcessed(survey = "all")

# Download only the essential remote spatial data files
getGeoProcessed(dataset = "essential")
```

You will be prompted to enter the password for the Google Drive account storing the remote files (password: fusethis!). The download will take a few minutes. The files are automatically placed in the appropriate sub-directories of `/fusionData`, with the directories created if necessary. After successful download, your fusionData "system" is ready to go.

## Usage and structure

Although fusionData is structured (and loadable) as a R package, it is better to think of it as a code and data *repository* that is shared and continuously modified by authorized users. fusionData is expected to grow over time as new surveys and spatial datasets -- and the code needed to process and manipulate them -- are added. fusionData is *not public*.

As you modify code and files in your local `/fusionData` project directory, you will need to commit and push those changes to the Github repository for them to be accessible to other users. In addition, it is good practice to pull the latest version of the repository from Github prior to making any modifications. That way, you know you are working from the latest shared version.

Since Github places limits on file/repository size, I decided to store certain data files "remotely" -- that is, outside of the Github repository. Specifically, the lowest-level "raw" data inputs and the associated "processed" versions of that data (details below). Over time, it is expected that these types of data files could become quite large in the aggregate.

The remotely-stored "raw" and "processed" data files are integral to the overall fusionData "system", but they are not present in the Github repository itself. Instead, the remote files (and associated directory structure) are stored in Google Drive and can be automatically and safely added to a user's *local* fusionData folder using provided functions. Once the remote files are added, the user's local fusionData package is fully functional.

Remote data files should be rather static over time. So, it is expected that a user will need to update (re-download) the remote files to their local fusionData directory only infrequently. However, when users add or modify *code* (or smaller, ancillary data files), the changes are then pushed to the Github repository where they become subject to code reviews, versioning control, and accessible to other authorized users.

In short: The Github repository stores any and all code needed to build and document the fusionData architecture. But certain, infrequently-modified data files are stored remotely. Users can add the remote files to their local installation of fusionData and only code and ancillary data files are pushed to Github when modified.

Changes or additions to the remote files is likely to be rare, and will be done "by hand" to prevent any inadvertent changes.

Below is an overview of the top-level directories in the fusionData repository, including both Github-based and "remote" elements.

### `/R`
`.R` scripts defining functions for doing "fusionData things". Not all are exported.

### `/data`
Shared, package-wide `.rda` data files. Loadable via `data()`, as usual for R packages.

### `/data-raw`
`.R` scripts needed to create any package-wide `.rda` objects in `/data`, as usual for R packages.

### `/man`
Documentation (i.e. "manual") of functions in `/R` and data in `/data`, as usual for R packages.

### `/universe`
Directory for the "Universal Survey Dictionary" Shiny app. The app itself can be run by calling `universe()`.

### `/harmony`
Directory for the "Survey Harmonization Tool" Shiny app. The app itself can be run by calling `harmony()`.

### `/survey-processed`
Contains processed survey data and associated code. Subdirectories refer to specific surveys and vintages.

Example: `/survey-processed/RECS/2015`

The Github repository version of `/survey-processed` contains two kinds of files:

1. Custom .R scripts that transform *raw* survey microdata (located in `/survey-raw`) into "processed" versions that adhere to certain requirements, structure, and naming conventions. 

Example: `/survey-processed/RECS/2015/RECS_2015_H_processed.R`

2. "Dictionary" files (.rds) that contain standardized metadata and variable descriptions for a particular survey. 

Example: `/survey-processed/RECS/2015/RECS_2015_H_dictionary.rds`

`/survey-processed` also includes .fst files containing the processed microdata itself. These files are stored remotely and can be added to a user's local fusionData directory by calling `getSurveyProcessed()`.

For example, the .fst file `/survey-processed/RECS/2015/RECS_2015_H_processed.fst` contains processed, household-level microdata for the 2015 RECS. The *code* that creates this file is found in `/survey-processed/RECS/2015/RECS_2015_H_processed.R`. The .R file is part of the Github repository ([see here](https://github.com/ummel/fusionData/blob/master/survey-processed/RECS/2015/RECS_2015_H_processed.R)), but the .fst file is stored remotely.

Use of [.fst files](http://www.fstpackage.org/) allows data to be read very quickly from disk, in part or in full. Functions in the fusionData package take advantage of this.

### `/survey-raw`
A remote directory (i.e. not present in the Github repository) containing *raw* survey data files. Subdirectories refer to specific surveys and vintages.

Example: `/survey-raw/RECS/2015`

`/survey-raw` can be downloaded and added to a user's local fusionData directory by calling `getSurveyRaw()`. However, in practice, there is no reason for a user to store raw survey data locally unless it is for a survey that they are actively processing or editing. Once the *processed* version of a survey (`*_processed.fst`) is stable and uploaded to the remote Google Drive, users can access and use the processed version without ever needing to download or look at the original/raw data.

### `/geo-processed`
Contains processed spatial data and associated code. Subdirectories refer to specific spatial datasets.

Example: `/geo-processed/EPA-SLD`

The Github repository version of `/geo-processed` contains the following kinds of files:

1. Custom `.R` scripts that transform *raw* spatial data (located in `/geo-raw`) into processed .rds files that meet certain requirements.

Example: `/geo-processed/EPA-SLD/epa-sld_v3_processed.R`

`/geo-processed` also includes .rds files containing the processed spatial data itself. These files are stored remotely and can be added to a user's local fusionData directory by calling `getGeoProcessed()`. 

For example, the .rds file `/geo-processed/EPA-SLD/epa-sld_v3_processed.rds` contains processed spatial variables from version 3 of the EPA's [Smart Location Database (SLD)](https://www.epa.gov/smartgrowth/smart-location-mapping#SLD). The *code* that creates this file is found in `/geo-processed/EPA-SLD/epa-sld_v3_processed.R`. The .R file is part of the Github repository ([see here](https://github.com/ummel/fusionData/blob/master/geo-processed/EPA-SLD/epa-sld_v3_processed.R)), but the .rds file is stored remotely.

Importantly, the `/geo-processed` remote content *also* includes three essential spatial data files that are, in practice, all that most users will need to perform data fusion locally. These files and their roles are described in more detail later on.

1. `geo_predictors.fst`
1. `concordance/geo_concordance.fst`
1. `concordance/bg_centroids.rds`

For users who are not modifying or adding spatial datasets, it is sufficient to call `getGeoProcessed(dataset = "essential")` to load the three essential "geo" files.

### `/geo-raw`
A remote directory (i.e. not present in the Github repository) containing *raw* spatial data files. Subdirectories refer to specific spatial datasets.

Example: `/geo-raw/EPA-SLD`

`/geo-raw` can be downloaded and added to a user's local fusionData directory by calling `getGeoRaw()`. However, in practice, there is no reason for a user to store raw spatial data locally unless it is for a spatial dataset that they are actively processing or editing.

## Ingest survey data

"Ingesting" a survey requires transforming raw survey data into "processed" (i.e. standardized) microdata files that meet certain requirements. The fusionData codebase depends on the processed microdata having recognizable structure and features.

The ingestion process for each survey is documented and defined by a .R script (possibly multiple scripts) that must be written manually. The goal is to produce a data.frame containing microdata observations that (ideally) meet the following conditions:

* Contains as many observations and variables as possible
* Variable names and descriptions are taken from the official codebook
* Codes used in the raw data are replaced with descriptive labels from the codebook
* All "valid blanks" in the raw data are set to plausible values
* All "invalid blanks" or missing values in the raw data are imputed
* Ordered factors are used and defined whenever possible (as opposed to unordered)
* Every column contains a variable description as a `labelled::var_label` attribute
* Standard names for unique household/person identifiers and observation weights (including replicate weights)
* Variables identifying respondent location are consistent with those defined in `geo-processed/geo_concordance.fst`

Let's look at a few of the variables in the processed RECS 2015 microdata to get a sense of what preferred output looks like. Note that the file name includes a `_H_` identifier, indicating that the microdata in question is household-level. Surveys that include both household and person-level respondent information have two such files -- both "H" and "P" microdata. The RECS has only household ("H") microdata.

```{r, echo = TRUE}
recs <- fst::read_fst("survey-processed/RECS/2015/RECS_2015_H_processed.fst")
head(select(recs, recs_2015_hid, weight, sizeofgarage, recs_iecc_zone, rep_1))
```
Notice that the household ID variable has a standardized name ("recs_2015_hid"), as does the observation weights column ("weight") and the first of the 96 replicate weights ("rep_1"). If the microdata consisted of person-level observations nested within households (e.g. as in the ACS), it would have have an additional "pid" integer variable to uniquely identify each person within the household.

The variable "recs_iecc_zone" tells us something about a respondent's location ([IECC climate zone](https://basc.pnnl.gov/images/iecc-climate-zone-map)). This and other spatially-referenced variables are defined and named to be consistent with variables in the `geo-processed/geo_concordance.fst` file (more on this file below). This allows subsequent operations to intelligently impute each respondent's location (see `?imputePUMA`).

The name of the other variable ("sizeofgarage") comes from the RECS codebook. In the case of "sizeofgarage", the raw data contained NA's (valid "skips") for households without a garage. Those blanks are replaced with an intelligible label ('No garage'). In addition, "sizeofgarage" is classed as an ordered factor, since the labels have a natural ordering.

```{r, echo = TRUE}
class(recs$sizeofgarage)
levels(recs$sizeofgarage)
```

You can see how, exactly, the raw data was transformed by viewing the associated code in `/survey-processed/RECS/2015/RECS_2015_H_processed.R`.

Given the variety of survey data structures and conventions, there is no strict procedure for how the .R file should be written. However, there are common steps and tools likely to be applicable to most surveys. The `RECS_2015_H_processed.R` script is a good "template" in this regard, since it includes many common operations -- including imputation of NA's using the provided `imputeMissing()` function.

The RECS 2015 has a comparatively simple microdata and documentation structure: household-level microdata in a single .csv file with an associated .xls codebook. Other surveys require more complex steps to assemble the necessary microdata. There is no limit on the number or nature of .R files that can be used to ingest a survey. If multiple .R files are used, the file names should include a two-digit sequence at the front to indicate the order in which the scripts are employed (`01*.R`, `02*.R`, etc.). The .R files located at [survey-processed/ACS/2019](https://github.com/ummel/fusionData/tree/master/survey-processed/ACS/2019) are an example of this.

*The .R files should include liberal use of comments to help others understand the code later. Good practice is for comments to explain **why** a piece of code is included, not just **what** it does.*

In all cases, the .R script that eventually saves the `_processed.fst` microdata file to disk must include the use of `labelled::set_variable_labels` to assign variable descriptions (ideally, taken from the official codebook) for each column. The same script must then call the `createDictionary()` function to create and save a standardized "dictionary.rds" file. You can see this at the end of [RECS_2015_H_processed.R](https://github.com/ummel/fusionData/blob/master/survey-processed/RECS/2015/RECS_2015_H_processed.R). Here is the resulting dictionary file for RECS 2015.

```{r, echo = TRUE}
recs.dictionary <- readRDS("survey-processed/RECS/2015/RECS_2015_H_dictionary.rds")
head(recs.dictionary)
```

In practice, there is no reason for a typical user to ever open a survey's dictionary file. The preferred and much more useful way to explore survey metadata and variable descriptions is described in the next section.

## Document variables

The previous section showed how and where a survey's "dictionary.rds" file(s) are created. Whenever a dictionary file is added or updated, it is necessary to run the `compileDictionary()` function to compile the individual survey dictionaries into a single "universal" dictionary. The usage is straightforward:

```{r, echo = TRUE}
compileDictionary()
```

As the console output tells us, `compileDictionary()` updates two files: `data/dictionary.rda` and `data/surveys.rda`. These files are part of the Github repository and are used by both the "Universal Survey Dictionary" and "Survey Harmonization Tool" that are part of fusionData.

The "Universal Survey Dictionary" is a Shiny app that can be accessed by the following call:

```r
# Open "Universal Survey Dictionary" Shiny app
universe()
```
This will open the app in a browser window. The tool allows the "universe" of available variables -- across all ingested surveys -- to be sorted and searched. A user should consult the universal dictionary after initial ingestion of a new survey, because it is an effective way to identify variables that need additional editing.

We may eventually make the app public so that potential fusionACS users can browse the universe of available fusion variables.

## Harmonize variables

The statistical linchpin of the fusion process is the set of "harmonized" variables common to a donor survey and the ACS. Identifying conceptually similar variables across surveys and determining how they can be modified to measure similar concepts is one of the most important steps in the process. It is also potentially time-consuming and error-prone.

The "Survey Harmonization Tool" was created to address these problems. It is a Shiny app that makes it easier to detect, create, and save "harmonies" among the variables of donor surveys and the ACS. The app launches in a browser window with the following call:

```r
# Open "Survey Harmonization Tool" Shiny app
harmony()
```

TO DO: Recording demonstrating app functionality.

When a user clicks "Submit harmony" within the app, the currently-specified "harmony" (as defined by the selected variables and settings) is saved to disk. Specifically, the details of that particular harmony are added to the appropriate .R "harmony file" located at `/harmony/harmonies`. For example, the file describing how to harmonize RECS 2015 and ACS 2019 variables is [RECS_2015__ACS_2019.R](https://github.com/ummel/fusionData/blob/master/harmony/harmonies/RECS_2015__ACS_2019.R). When `dget`-d, harmony files return a list of lists, where each element defines a harmony. Like this one, defining the harmony between the "fuelheat" variable in the RECS and the "hfl" variable in the ACS.

```r

  fuelheat__hfl = list(
    RECS = list(
      groups = 1:7,
      levels = c("Do not use space heating", "Electricity", "Fuel oil/kerosene", "Natural gas from underground pipes", "Propane (bottled gas)", "Some other fuel", "Wood (cordwood or pellets)"),
      breaks = "",
      adj = ""),
    ACS = list(
      groups = c(5, 6, 2, 3, 1, 6, 6, 4, 7),
      levels = c("Bottled, tank, or LP gas", "Coal or coke", "Electricity", "Fuel oil, kerosene, etc.", "No fuel used", "Other fuel", "Solar energy", "Utility gas", "Wood"),
      breaks = "",
      adj = ""),
    ordered = FALSE,
    comment = "",
    modified = "2021-06-10 10:00:33"),

```
This list object contains all of the information necessary to construct RECS and ACS microdata containing a new variable called "fuelheat__hfl" -- the harmonized version of the two associated heating fuel variables. This is precisely what the `harmonize()` function does (typically when called by `prepare()` as explained below) using all of the harmonies available in the specified harmony file. Note that harmonized variables are always indicated by a double-underscore ("__").

This strategy -- using the `harmony()` app to manually define harmonies and then letting `harmonize()` take care of subsequent data manipulation -- makes the construction of harmonized microdata easier, faster, and *much* safer.

Most users will eventually find themselves constructing harmonies via the app. Submitted harmonies are saved to *local* harmony files, which means you must commit and push those changes for them to show up in the Github repository -- and become available for others to use. This also means it is important to pull the most recent version of the repository when you begin working with fusionData. Otherwise, you risk duplicating the efforts of someone else.

## Compile spatial data

fusionData allows for spatially-referenced data to be merged with survey microdata, thereby expanding the set of potential predictor variables available in the fusion process. The geographic "unit of analysis" in this case consists of [PUMA's](https://www.census.gov/programs-surveys/geography/guidance/geo-areas/pumas.html), which are observed for ACS households and can be imputed (see `?imputePUMA`) for donor households.

Ingestion of spatial datasets is generally less onerous than for survey data; there are fewer requirements that the processed data must meet. The general strategy will look familiar: Raw spatial data is stored in `/geo-raw`. The raw data is transformed to a "*_processed.rds" file stored in `/geo-processed`. The associated .R file is stored in the same location.

A processed spatial .rds file has only two hard requirements it must meet.

1. It must contain a "vintage" column indicating the time period of each observation. The vintage can be a year, a year range ("2015-2016"), or the special value "always". The "always" value indicates that a measurement is time-invariant (e.g. a long-term climate "normal").
1. It must contain a column (or columns) whose name and values are also found in the "geo_concordance.fst" file.

Ordered factor variables should be classed as such; other categorical variables can be character. It is not (currently) necessary to document the variables, name them a certain way, or create a dictionary. Let's look at an example.

```{r, echo = TRUE}
irs <- readRDS("geo-processed/IRS-SOI/IRS-SOI_2018_processed.rds")
head(irs[, 1:5])
```
The `irs` object contains processed spatial data constructed from the [IRS Statistics of Income (SOI)](https://www.irs.gov/statistics/soi-tax-stats-individual-income-tax-statistics-zip-code-data-soi) zip code tax return data for 2018. The underlying raw data is stored remotely at `/geo-raw/IRS-SOI/2018`. The script used to create the "processed.rds" file is [available here](https://github.com/ummel/fusionData/blob/master/geo-processed/IRS-SOI/IRS-SOI_AllVintages_processed.R). The "zcta10" column indicates the Zip Code Tabulation Area (circa 2010) associated with each observation.

The "zcta10" variable is also found in the "geo_concordance.fst" file, which contains information about how to link geographic units to PUMA's. Its creation relies heavily on data from the Missouri Census Data Center's [Geocorr engine](https://mcdc.missouri.edu/applications/geocorr.html) (code [here](https://github.com/ummel/fusionData/blob/master/geo-processed/concordance/02%20geo_concordance.R)). The information on how to link zip codes to PUMA's is used to aggregate the IRS-SOI data to PUMA-level prior to merging with survey microdata.

The "geo_concordance.fst" file contains a variety of variables that can be used to identify the location of observations in a processed spatial data file. Most of these are documented by [Geocorr](https://mcdc.missouri.edu/applications/docs/maggot2014.html). Others were added within the `geo-processed/concordance/02 geo_concordance.R` file to allow concordance with variables found in particular datasets. The concordance file can be expanded over time as necessary.

```{r, echo = TRUE}
concordance <- fst::fst("geo-processed/concordance/geo_concordance.fst")
names(concordance)
```
In some cases, the processed .rds file will include multiple variables to achieve concordance. For example, a spatial dataset with block group observations must include columns for "state", "county10", "tract10", and "bg10" in order to allow a smooth merge with the concordance file (this is the case for the EPA-SLD dataset).

Unlike with processed *survey* data, the naming convention for processed spatial data files is quite relaxed. The function `compileSpatial()` automatically detects and compiles all files in `/geo-processed` ending with "_processed.rds". As long as a processed spatial data file has the necessary suffix -- and meets the two requirements above -- it will be compiled into the `geo_predictors.fst` file. Whenever a processed .rds file is added or updated, it is necessary to run `compileSpatial()` to update the `geo_predictors.fst` file.

The `geo_predictors.fst` file contains all variables and vintages across available spatial datasets, aggregated to PUMA-level in preparation for merging with survey microdata. The structure of this file is unusual, but it is not intended to be worked with directly. It is designed to allow the `prepare()` function (demonstrated below) to efficiently read the necessary data from disk when it merges spatial variables for particular donor and recipient surveys.

Consequently, unless a user is actively adding or editing processed spatial data, the only "geo files" that are strictly necessary for the fusion process are `geo_predictors.rds` and `geo_concordance.rds`, both of which can be obtained by calling `getGeoProcessed(dataset = "essential")`.

## Prepare for fusion

The following example shows how the fusionData `prepare()` function is used to assemble complete, consistent, and harmonized microdata that can then be passed to the [fusionModel package](https://github.com/ummel/fusionModel) to fuse donor variables to ACS microdata.

The simplest usage is shown below. In this case, we are requesting microdata outputs at the household-level that will allow us to (subsequently) fuse RECS 2015 donor variables to ACS 2019 recipient microdata.

```{r, echo = TRUE, results = 'hide'}
# Prepare RECS 2015 household microdata for fusion with ACS 2019 microdata
data <- prepare(donor = "RECS_2015", recipient = "ACS_2019", respondent = "household")
```
The resulting `data` object is a list containing two data frames. The first slot contains the "prepared" donor microdata. The second slot contains the "prepared" ACS recipient microdata. Notice that the RECS microdata has many more variables/columns than the ACS data. This is because `prepare` donor output includes -- by default -- both the "harmonized" variables as well as any other donor variables not used to create harmonies. The latter are potential candidates for fusion (and there are many in the case of RECS).

```{r, echo = TRUE}
lapply(data, dim)
```
A key purpose of `prepare()` is to harmonize the donor and recipient "shared" variables. This is done internally by the `harmonize()` function, using the variable harmonies created by users via the `harmony()` tool. In this case, `harmonize()` is using the [RECS_2015__ACS_2019.R](https://github.com/ummel/fusionData/blob/master/harmony/harmonies/RECS_2015__ACS_2019.R) file to harmonize the donor and recipient microdata. Let's confirm that a few of the shared variables are, indeed, harmonized.

```{r, echo = TRUE}
v <- names(data$ACS_2019)[2:6]
head(data$RECS_2015[v])
head(data$ACS_2019[v])
```
Notice that the harmonized variable *values* are typically integers (possibly factorized); that is, they contain no intelligible labels. This is because `harmonize()` maps each original value/level to a (integer) group assignment as specified in the relevant `.R` harmony file. The one exception is when *numeric* variables in the two surveys are conceptually identical and can be included without modification.

Since RECS and ACS are both nationally representative surveys, the distribution of the harmonized variables should look pretty similar across the two data frames. We can confirm this for the `fuelheat__hfl` variable, which creates harmony between the RECS and ACS heating fuel variables ("fuelheat" and "hfl", respectively). We can compare the proportion of cases by harmonized value.

```{r, echo = TRUE}
round(table(data$RECS_2015$fuelheat__hfl) / nrow(data[[1]]), 3)
round(table(data$ACS_2019$fuelheat__hfl) / nrow(data[[2]]), 3)
```
`prepare()` also merges spatial variables with both the donor and recipient microdata. The function `imputePUMA()` is used internally to randomly assign a plausible PUMA to each donor household. Pre-compiled spatial variables (those in `geo-processed/geo_predictors.fst`) are then merged onto both the donor and recipient microdata at the PUMA level. Spatial variables are indicated by the double-dot ("..") in the variable name, analogous to the way that harmonized variables are indicated by the double-underscore ("__"). 

Let's look at the variables in the recipient ACS microdata.

```{r, echo = TRUE}
names(data$ACS_2019)
```
The string to the left of the ".." identifies the spatial dataset; the string to the right is a unique, syntactically-valid identifier. It's not critical that specific spatial variables be identifiable in the fusion process. And because pre-processing of spatial datasets does not impose a stringent naming/documentation convention (by design), these non-nonsensical-but-unique names are the safest way to identify spatial variables.

Now let's explore some of the additional arguments to `prepare()`:

1. We can pass unquoted donor variable names and/or selectize statements to the `...` argument if we want to return a specific set of fusion variables for the donor (instead of all variables, the default behavior). This is passed internally to `completeDonor()`.

1. The `implicates` argument controls how many PUMA's are imputed for each donor household. Setting `implicates` higher results in more variability in the spatial predictors merged to a given household, reflecting our uncertainty about where the household is located. The use of implicates here mimics usage in standard multiple imputation techniques (5 implicates is typical).

1. We can limit the spatial datasets merged to the microdata via the `spatial.datasets` argument. Default is to include all available datasets, and this is sensible in most cases.

1. The `window` argument controls how wide a timespan we tolerate when merging spatial variables to microdata. The default (`window = 0`) means that spatial variables are merged only when their vintage matches that of the microdata. A larger `window` will generally mean more spatial variables in the output but at some cost in terms of temporal alignment.

1. The `pca` argument controls whether/how principal components analysis (PCA) is used to reduce dimensionality of the spatial variables. `?prepare` provides additional details concerning the `pca` argument.

The following shows a more complex (and realistic) call to `prepare()`, making use of the optional arguments.

```{r, echo = TRUE, results = 'hide'}
# Prepare RECS 2015 household microdata for fusion with ACS 2019 microdata
data <- prepare(donor = "RECS_2015", 
                recipient = "ACS_2019", 
                respondent = "household",
                cooltype, agecenac, kwhcol,  # Request specific donor variables
                implicates = 5,
                window = 3,
                pca = c(30, 0.9))
```
```{r, echo = TRUE}
lapply(data, dim)
```
The number of observations in the donor microdata is now higher, reflecting the use of `implicates = 5`. Note that the number of rows has increased by less than a factor of five. This is because `imputePUMA()` collapses duplicate household-PUMA observations and adjusts the sample "weight" column accordingly. This reduces the amount of data without affecting subsequent statistical results.

There are now more spatial variables merged to both the donor and recipient due to setting `window = 3`. There would be even more spatial variables if we had not specified the `pca` argument. Doing so collapsed the *numeric* spatial variables into a smaller number of components -- now identified by the "pca.." prefix -- which we can confirm by looking at the recipient column names.

```{r, echo = TRUE}
names(data$ACS_2019)
```
The difference in the number of columns between the RECS and ACS microdata is due to the former's inclusion of our three requested donor variables plus an observation "weight" column. Otherwise, both data frames are entirely consistent with one another. They each include a unique household identifier variable and an identical set of harmonized survey and spatial variables that can be exploited by the fusion process.

## Make it rain

At this point, we are ready to fuse. This is straightforward using the `train()` and `fuse()` functions from the fusionModel package. To make this even easier, `prepare()` donor output includes a "fusion.vars" attribute that can be passed directly to `train()` to identify the variables available to be fused.

```{r, echo = TRUE, results = "hide"}
fit <- fusionModel::train(data = data$RECS_2015, 
                          y = attr(data$RECS_2015, "fusion.vars"), 
                          ignore = "recs_2015_hid", 
                          weight = "weight",
                          mc = TRUE,
                          maxcats = 10,
                          lasso = 0.95)
```

We then `fuse()` (i.e. simulate) the fusion variables onto the harmonized ACS microdata. This is a non-trivial exercise, since the recipient ACS microdata has `r nrow(data$ACS_2019) ` observations. Setting `induce = FALSE` eases the computation and memory burden considerably. The call below shouldn't require more than about 5GB of RAM.

```{r, echo = TRUE, results = "hide"}
sim <- fusionModel::fuse(data = data$ACS_2019, 
                         train.object = fit, 
                         induce = FALSE)
```

A quick check that the fusion output looks plausible:

```{r, echo = TRUE}
nrow(sim)
head(sim)
summary(data$RECS_2015$kwhcol)
summary(sim$kwhcol)
```

Onward and upward!
