---
title: "fusionData"
author: Kevin Ummel ([ummel@sas.upenn.edu](mailto:ummel@sas.upenn.edu))
output:
  md_document:
    variant: gfm
#output: html_document  # Useful for testing/editing; turned off for final production
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      #cache = TRUE, # Useful for testing/editing; can be turned off for final production
                      comment = NA)
```

```{r, echo = FALSE, include = FALSE}
library(fusionData)
```

## Overview

**fusionData** is used to create and manage the data inputs underpinning the larger fusionACS platform. It has a number of objectives:

1. Process raw survey microdata using a standard approach and format.
1. Consistently document survey variables to create a "universal" data dictionary.
1. Harmonize variables in "donor" surveys with those in the American Community Survey (ACS).
1. Create harmonized donor and ACS microdata to be used by [fusionModel](https://github.com/ummel/fusionModel).
1. Process and integrate spatially-referenced data available to the fusion process.

Each of these objectives is described in more detail below along with example usage.

## Structure and usage

Although fusionData is structured (and loadable) as a R package, it is better to think of it as a code and data *repository* that is shared and continuously modified by authorized users. fusionData is expected to grow over time as new surveys and spatial datasets -- and the code needed to process and manipulate them -- are added. fusionData is *not public*.

Since Github places limits on file/repository size, I decided to store certain data files "remotely" -- that is, outside of the Github repository. Specifically, the lowest-level "raw" data inputs and the associated "processed" versions of that data (details below). Over time, it is expected that these types of data files could become quite large in the aggregate.

The remotely-stored "raw" and "processed" data files are integral to the overall fusionData "system", but they are not present in the Github repository itself. Instead, the remote files (and associated directory structure) are stored in Google Drive and can be automatically and safely added to a user's *local* fusionData folder using provided functions. Once the remote files are added, the user's local fusionData package is fully functional.

Remote data files should be rather static over time. So, it is expected that a user will need to update (re-download) the remote files to their local fusionData directory only infrequently. However, when users add or modify *code* (or smaller, ancillary data files), the changes are then pushed to the Github repository where they become subject to code reviews, versioning control, and accessible to other authorized users.

In short: The Github repository stores any and all code needed to build and document the fusionData architecture. But certain, infrequently-modified data files are stored remotely. Users can add the remote files to their local installation of fusionData and only code and ancillary data files are pushed to Github when modified.

Changes or additions to the remote files is likely to be rare, and will be done "by hand" to prevent any inadvertent changes.

Below is an overview of the top-level directories in the fusionData repository, including both Github-based and "remote" elements.

### `/R`
`.R` scripts defining functions for doing "fusionData things". Not all are exported.

### `/data`
Shared, package-wide `.rda` data files. Loadable via `data()`, as usual for R packages.

### `/data-raw`
`.R` scripts needed to create any package-wide `.rda` objects in `/data`, as usual for R packages.

### `/man`
Documentation (i.e. "manual") of functions in `/R` and data in `/data`, as usual for R packages.

### `/universe`
Directory for the "Universal Survey Dictionary" Shiny app. The app itself can be run by calling `universe()`.

### `/harmony`
Directory for the "Survey Harmonization Tool" Shiny app. The app itself can be run by calling `harmony()`.

### `/survey-raw`
A remote directory (i.e. not present in the Github repository) containing *raw* survey data files. Sub-directories refer to specific surveys and vintages (e.g. `/survey-raw/RECS/2015`). `/survey-raw` can be downloaded and added to a user's local fusionData directory by calling `getSurveyRaw()`.

### `/survey-processed`
Contains subdirectories that refer to specific surveys and vintages (e.g. `/survey-processed/RECS/2015`). The Github repository version of `/survey-processed` contains two kinds of files:

1. Custom `.R` scripts that transform *raw* survey microdata (located in `/survey-raw`) into "processed" versions that adhere to certain requirements, structure, and naming conventions. Example: `/survey-processed/RECS/2015/RECS_2015_H_processed.R`

1. "Dictionary" files (`.rds`) that contain standardized metadata and variable descriptions for a particular survey. Example: `/survey-processed/RECS/2015/RECS_2015_H_dictionary.rds`

Importantly, `/survey-processed` also contains `.fst` files with the processed microdata. Example: `/survey-processed/RECS/2015/RECS_2015_H_processed.fst`. These files are stored remotely. Use of [.fst files](http://www.fstpackage.org/) allows this data to be read very quickly from disk, in part or in full. The remotely-stored `*_processed.fst` files can be added to a user's local fusionData directory by calling `getSurveyProcessed()`.

<!--
### `/spatial-raw`
In development. Analogous to `survey-raw` but for spatial data.

### `/spatial-processed`
In development. Analogous to `survey-processed` but for spatial data.
--> 

## Process raw survey microdata

To do.

## Document survey variables

To do.

```r
# Open "Universal Survey Dictionary" Shiny app
universe()
```

## Harmonize variables across surveys

To do.

```r
# Open "Survey Harmonization Tool" Shiny app
harmony()
```

## Create harmonized microdata

To do.

## Integrate spatial data

To do.

## Putting it all together

The following example shows how fusionData functions can be used to assemble harmonized microdata that is then passed to the [fusionModel package](https://github.com/ummel/fusionModel) to fuse donor variables to ACS microdata.

We begin by using `harmonize()` to parse the information in a `.R` "harmony file" and create microdata containing the harmonized (i.e. shared) variables among the donor and recipient -- in this case, the RECS and ACS. The harmony file is produced using the Survey Harmonization Tool (`?harmony`). See `?harmonize` for details.

```{r, echo = TRUE, results = 'hide'}
# Harmonize RECS 2015 and ACS 2019 microdata
hdata <- harmonize(harmony.file = "RECS_2015__ACS_2019.R", respondent = "household")
```

The resulting `hdata` object is a list containing two data frames.

```{r, echo = TRUE}
lapply(hdata, dim)
```

We can look at the microdata in each of the data frames to confirm that the variables are, indeed, harmonized. Note that the harmonized variable *values* are typically integers (possibly factorized); that is, they contain no intelligible labels. This is because `harmonize()` maps each original value/level to a (integer) group assignment as specified in the `.R` harmony file. The one exception is when *numeric* variables in the two surveys are conceptually identical and can be included without modification.

```{r, echo = TRUE}
head(hdata$RECS_2015[, 1:5])
head(hdata$ACS_2019[, 1:5])
```

Since RECS and ACS are both nationally representative surveys, the distribution of the harmonized variables should look pretty similar across the two data frames. Let's confirm this for the `fuelheat__hfl` variable, which creates harmony between the RECS and ACS heating fuel variables ("fuelheat" and "hfl", respectively, in the original data).

```{r, echo = TRUE}
round(table(hdata$RECS_2015$fuelheat__hfl) / nrow(hdata[[1]]), 3)
round(table(hdata$ACS_2019$fuelheat__hfl) / nrow(hdata[[2]]), 3)
```

Looks good. The next step is to add "fusion variables" to the harmonized RECS microdata to create our "donor" dataset. The fusion variables are those we want to "fuse" from RECS to the ACS. The `completeDonor()` function provides a convenient way to do this. By default, it will add *all* valid fusion variables; i.e. variables not already used for harmonization. But we can also use `dplyr::select()`-style arguments to return only a subset of RECS variables. Like this:

```{r, echo = TRUE}
donor <- completeDonor(data = hdata$RECS_2015, cooltype, agecenac, kwhcol)
```

The resulting `donor` data frame contains all of the harmonized variables as well as the specified fusion variables.

```{r, echo = TRUE}
head(donor)
```

At this point, we are ready to fuse! This is straightforward using the `train()` and `fuse()` functions from the fusionModel package. To make this even easier, `completeDonor()` output includes a "fusion.vars" attribute that can be passed directly to `train()` to identify the variables to be fused.

```{r, echo = TRUE, results = "hide"}
fit <- fusionModel::train(data = donor, 
                          y = attr(donor, "fusion.vars"), 
                          ignore = "recs_2015_hid", 
                          weight = "weight")
```

We then `fuse()` (i.e. simulate) the fusion variables onto the harmonized ACS microdata. This is a non-trivial exercise, since the recipient ACS microdata has `r nrow(hdata$ACS_2019) ` observations.

```{r, echo = TRUE, results = "hide"}
sim <- fusionModel::fuse(data = hdata$ACS_2019, train.object = fit)
```

A quick check that everything looks plausible:

```{r, echo = TRUE}
nrow(sim)
head(sim)
summary(donor$kwhcol)
summary(sim$kwhcol)
```

Onward and upward!
